Alleviating the Topology Mismatch Problem in Unstructured and Structured Overlay Networks: A Survey

### 2014-01-27 unstructured.tex
2012.10.20 (Vassilis) Images reworked. Text alligned with the new images.

\begin{figure}[ht]
\centering
\subfigure[Iterative deepening with three levels.] {
  \includegraphics[scale=0.4]{img/algorithms/iterative_deepening}
  \label{figure:dbfs:iterdeep}
}\qquad\qquad
\subfigure[Directed BFS.] {
  \includegraphics[scale=0.4]{img/algorithms/directed_bfs}
  \label{figure:dbfs:dbfs}
}\qquad\qquad
\subfigure[Local indices with radius size equal to $2$.] {
  \includegraphics[scale=0.4]{img/algorithms/local_indices}
  \label{figure:dbfs:localindx}
}
\caption{Improving search in P2P networks}
\label{figure:dbfs}
\end{figure}

@@

In iterative deepening (see Figure~\ref{figure:dbfs:iterdeep}), the search is
performed on a BFS tree with multiple preset depths. The depth limit is
iteratively increased by the source node for each query, based on the quality of
the results. The source node may issue a new request with increased depth limit
that will trigger the nodes at the last depth level to resume the search. The
iterative approach avoids restarting the whole search process from scratch at
each iteration and reduces the load of the nodes on the upper levels of the
tree. Its major drawback is the delay between successive iterations, as the
source node needs to examine the results at each iteration before deciding to
quit or resume the query.

The directed BFS, as shown in Figure~\ref{figure:dbfs:dbfs}, tries to avoid this
delay by forwarding the query messages only to a selected set of neighbours, in
which the selection criteria varies from the number of results received
previously, distance in terms of hops, bandwidth, or the query load of the
neighbour.

In local indices, as shown in Figure~\ref{figure:dbfs:localindx}, each node
indexes data within a radius of $r$ hops and uses this local index to answer
queries on behalf of them without generating additional traffic. Local indices
greatly reduce the aggregate bandwidth usage of the network and improves query
efficiency, however, updating the indices in cases with frequent node joins and
leaves introduces a serious overhead to the system if the radius is kept broad.

@@

@@

\begin{figure}[ht]
\centering
\subfigure[Underlying network with edge costs.] {
  \includegraphics[scale=0.4]{img/algorithms/narada}
  \label{figure:narada:underlying}
}\qquad\qquad
\subfigure[Peer A sends a message to the rest using regular broadcast.] {
  \includegraphics[scale=0.4]{img/algorithms/narada2}
  \label{figure:narada:regu}
}\qquad\qquad
\subfigure[Peer A sends a message to the rest using end system multicast.] {
  \includegraphics[scale=0.4]{img/algorithms/narada3}
  \label{figure:narada:multicast}
}
\caption{A visualisation of the Narada protocol.}
\label{figure:narada}
\end{figure}

\emph{Narada} \cite{CRZ2000,CRSZ2001,CRSZ2002} is a generic
protocol for creating self-adapting overlay networks, that can achieve
application-layer multicast communication without requiring IP multicast
infrastructure at the network layer\footnote{
  IP multicast is the term reffering to the method of sending IP datagrams to a
  group of receivers in a single transmission. Even though the method is
  available some years now, IP multicast has not taken off as anticipated since
  many believe that it violates the stateless design of the current Internet.
  For this reason it is widely deployed in contained environments like
  enterprises (i.e. IPTV applications - distance learning, video conferencing),
  commercial stock exchanges, and multimedia content delivery networks.
}. Although it was not originally designed as a P2P system nor can be considered
scalable in the contemporary sense of the term, Narada has been pioneering
(maybe along with \emph{Scattercast} \cite{C2000}) in that
it was the first to consider the feasibility of overlay-based, application
layer, services over the Internet that can take into account bandwidth and
latency properties of the physical underlying infrastructure. Additionally,
authors realised the inefficiency caused by the topology mismatch problem and
tried to encounter it by building a richer connected graph, called a
\emph{mesh}, and building per source minimum spanning trees\footnote{
  ENarada \cite{LYL2008} used Gossip protocol for the
construction
}
as shown in Figure~\ref{figure:narada}. They also keep the graph and the trees
dynamically updated while nodes continue to join and leave the network.
Moreover, the protocol tries to ease the physical link stress, the overall
resource usage and the relative delay among end systems. Unfortunately, the most
important limitation of Narada is that although it works reasonably well for
small groups, it does not scale well for larger networks, therefore it is not
suitable for, potentially very large, P2P file sharing application networks.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Landmark Binning (LM)}\label{sec:landmark_binning}
\begin{figure*}
\centering
  \includegraphics[scale=0.4]{img/algorithms/landmark_binning}
\caption{Example 2D coordinate overlay and a sample routing path from node D to
(x,y).}
\label{fig:landmark_binning}
\end{figure*}

Landmark Binning \cite{RHMKS2002} is an approach to partition close
by nodes into bins based on their distance to well known anchor nodes across the
Internet, as shown in Figure~\ref{fig:landmark_binning}. In order to detect
locality, nodes use network latency (i.e. RTT) as a measurement technique. The
network latency, even though not always accurate, is selected in this work due
to its non-intrusive, transparent and easy to apply behaviour. In order for the
binning to work, few well known anchor servers with known physical locations on
the Internet. Authors estimate that a number of 12 such servers can prove
sufficient for this. Every newly arriving node measures its distance from these
landmarks and unilaterally decides to join a specific bin based on these
results. Specifically, the node measures its round-trip time to each of the
landmarks and orders the values in a decreasing order. The ordering represents
a \emph{bin}, in the sence of close-by nodes having the same landmark ordering
and hance belong to the same bin. This means that a landmark system consisting
of $m$ such nodes results in $m!$ potentional different bins.The operation of
the method is independent of the model incorporated by the overlay network and
thus it can be applied with no significant changes to both structured and
unstructured P2P systems. \cite{RHMKS2002} has the detailed description of such
algorithm for both architectures. Landmark binning is also considered as a good
candidate for content distribution networks in particular. The major
disadvantage of landmark binning is to install and maintain landmark servers on
different autonomous systems, all over the world. A typical P2P network usually
has a couple of million nodes connected at any given time, thus rendering
landmark servers an important parameter when designing for large scales.  The
authors claim that latency estimation does not drain much from the network
resources but in case scalability problems arise, they claim the solution is
replacing single landmark servers with clusters of servers within the same
physical area. However, this approach does not reduce the possibility of
excessive network traffic flow through these landmark servers. One other
possible problem is the incorrect binning caused by the use of inaccurate
methods for delay measurement, since tracking down, for example, network
latencies is far from proven to be safe for location estimation.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Topology adaptation}
Researchers initially started considering alternatives for services who
implement one-to-many communication schemes, like IP multicast, on the
application layer (i.e. Narada). Application level multicast protocols
are generic protocols that can be applied to both P2P file sharing systems, as
well as content distribution networks. Application layer multicast, however, is
not as efficient as the one enabled by IP and still struggles to ease the
difficulties imposed by the topology mismatch problem.

The topology adaptation based protocols modify the topology of the P2P network
using various techniques. The two commonly used approaches that have been
investigated in this section are creating \emph{spanning tree}s using connection
graphs, and creating \emph{clusters} of physically close nodes.

Narada and subsequent algorithms like AOTO, LTM, SBO, try to the problem by
building a richer connected graph and forming minimum spanning trees over this
graph that can efficiently route messages among peers. Even though spanning
trees provide efficient query performance, the construction and the update costs
of the spanning tree, especially in dynamic environments with many nodes joining
and leaving the network, cause large traffic overhead to the corresponding
network \cite{CRZ2000,CRSZ2001,CRSZ2002}. Thus, the original Narada algorithm is
designed for small groups therefore it is not scalable. Its followers try to
solve the scalability problem by introducing various methods including forming
minimum spanning trees for the two-hop neighbours ($N^2$) of each node,
partitioning the graph into two random groups where each group is responsible of
different tasks, or performing local optimizations dynamically on the overlay
graph. The advantage of building minimum spanning trees is that they maintain
the connectivity on the network in an efficient manner, while still preserving
the overall search scope. However, building and maintaining a minimum spanning
tree creates additional overhead for the network.

The cluster based approaches, on the other hand, select to link physically
closer nodes with each other. T2MC, for example, uses traceroute logs and DDNO
uses domain names to cluster close by nodes. Unfortunatelly, commonly used
methods for proximity detection across Internet do not always return
reliable results, therefore mapping accuracy is not guaranteed. Moreover,
traceroute for example, is a heavy weight utility to be frequently used on the
network, let alone the fact that many network equipment vendors do not allow
traceroute calls at all. The most problematic aspect of clusterting, though, is
its nature per se. Limited connectivity among the various local domains can
significantly shrink the search scope, negatively affecting the query response
time the P2P end-user experiences. Of course, techniques that try to round up
such edges and balance the efficiency of the clustering approach with improved
connectivity, have also been proposed, like DDNO, which attempts to address this
limitation by forcing half of each node's connections to be with other, randomly
selected, nodes over the network

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Forwarding optimisation}
In the first realisation of the Gnutella protocol, each query request received
by a peer was then forwarded to all of its neighbours thus generating
unnecessary traffic over the network. Forwarding-optimisation-based approaches,
propose intelligent neighbour selection for a message's next hop across the
routing path, in order to replace that inefficient \emph{blind flooding} scheme
of Gnutella's. The selection criteria vary depending on the algorithm, but they
mainly use one or multiple statistical metrics that span from some obvious ones,
like the latency of the link between the nodes or the exposure of specific
features such as high processing bandwidth, high capacity, or low latency, to
more sophisticated alternatives that may even take into account application
level requirements such as the reliability of the neighbour.
These appoaches have the advantage of enhancing the search responsiveness but on
the other hand suffer from the drastic reduction of the search scope thus
limiting the scalability of the whole network. This, in practice reduces the
aggregate resource usage of the P2P network and improves the overall performance
of the system. On the other hand, the algorithms do not offer a real solution to
the topology mismatch problem. Thus, none can guarantee that overlay and
underlying topologies are aligned with each other let alone quantify the
mismatch and try to alleviate. Broadcast-based methods can be used in
conjunction with other approaches to improve the quality of the P2P systems.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Caching and replication}
Caching is a widely used technique to exploit locality and minimise redundant
transfer of data, initially and successfully adopted by web and file server
application environments. Since peers in a P2P system also operate as
servers, it is intuitively expected that P2P file sharing systems can also
benefit from caching in improving performance and reducing overall resource
usage. However, the design of caches in this context is non-trivial compared to
the web-based caching. Exactly, because each node can act as both a server and a
client, two important issues have to be considered at design time. First, the
lifetime of a query is short, as the nodes join and leave frequently. Second,
the result of a single query string is not always the same, as it is dependent
on the source of the query, the TTL value set for the messages, the current
interconnection of peers and the high volatility of the environment. So, in
order to develop a successful caching system for a P2P architecture, these
parameters also have to be considered. P2P caching/replication can be applied
on two different levels, namely caching indices or pointers to data or caching
the data itself. Successfull implementations have already been developed in
some commercial P2P systems, like KazaA, which uses caching of indices on its
super-peer level. Unfortunatelly, even though the state of the art in P2P
algorithms using caching methods reduce the resource usage of the network, they
also fail to consistently address the probelem of the topology mismatch.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Landmarking}\label{sec:landmark}
In landmark-based algorithms, nodes use network delay (e.g. RTT) as a
distance measurement method to position themselves with respect to ``a priori''
known servers (the landmark servers) on the Internet. Landmark servers are used
by nodes to estimate their positions based on the intutive assumption that nodes
with similar distances to a set of landmarks, are physically close to each
other, as well, over the network. The landmark based protocol has two important
drawbacks. Firstly the network delay is not a reliable distance estimation
method. For example, based on the load on the network the delay to certain nodes
or networks can change from time to time, which will eventually affect the
distance measurements and wrong measurements will lead to wrong estimated
positions for the nodes, or incorrect and non optimal clusterings of the nodes.
The second drawback of using landmark servers is the cost of installing and
maintaining this kind of infrastructure across the whole Internet and for all
the different autonomous system domains. As popular P2P file sharing
applications usually have millions of peers connected at any time, it is not
false to assume that the network costs of maintaining these landmark servers
will be quite high. A possible solution to the scalability problem of the static
landmark servers is to use ordinary nodes as dynamic landmarks once they
estimate their own positions. Even though this approach scales much better than
static landmark servers, still the measurement accuracy problem affects the
overall performance of the system. Moreover,it can be characterised as a rather
coarse-grained approximation, therefore not particularly well suited for
detecting the correct positions of nodes within close distance to each other.

>>>>

% \begin{figure}[ht]
% \centering
% \subfigure[Iterative deepening with three levels.] {
%   \includegraphics[scale=0.4]{img/algorithms/iterative_deepening}
%   \label{figure:dbfs:iterdeep}
% }\qquad\qquad
% \subfigure[Directed BFS.] {
%   \includegraphics[scale=0.4]{img/algorithms/directed_bfs}
%   \label{figure:dbfs:dbfs}
% }\qquad\qquad
% \subfigure[Local indices with radius size equal to $2$.] {
%   \includegraphics[scale=0.4]{img/algorithms/local_indices}
%   \label{figure:dbfs:localindx}
% }
% \caption{Improving search in P2P networks}
% \label{figure:dbfs}
% \end{figure}

@@

In iterative deepening, %(see Figure~\ref{figure:dbfs:iterdeep})
the search is
performed on a BFS tree with multiple preset depths. The depth limit is
iteratively increased by the source node for each query, based on the quality of
the results. The source node may issue a new request with increased depth limit
that will trigger the nodes at the last depth level to resume the search. The
iterative approach avoids restarting the whole search process from scratch at
each iteration and reduces the load of the nodes on the upper levels of the
tree. Its major drawback is the delay between successive iterations, as the
source node needs to examine the results at each iteration before deciding to
quit or resume the query.

The directed BFS%, as shown in Figure~\ref{figure:dbfs:dbfs},
tries to avoid this delay by forwarding the query messages only to a selected
set of neighbours, in which the selection criteria varies from the number of
results received previously, distance in terms of hops, bandwidth, or the query
load of the neighbour.

In local indices%, as shown in Figure~\ref{figure:dbfs:localindx},
each node
indexes data within a radius of $r$ hops and uses this local index to answer
queries on behalf of them without generating additional traffic. Local indices
greatly reduce the aggregate bandwidth usage of the network and improves query
efficiency, however, updating the indices in cases with frequent node joins and
leaves introduces a serious overhead to the system if the radius is kept broad.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph*{ \bf Delay Aware P2P System}
A new \emph{Delay Aware P2P System (DAPS)} is introduced in \cite{ZL2005}. Its
main goal is to reduce the time of a look-up request by dividing the routing
tables of peers into several sectors of increasing delay. The source node
emitting the query message defines a delay boundary or the \emph{pruning
factor} in the paper's context. Request messages are then forwarded only to
nodes whose delay is less than or equal that boundary. With the clustered
routing tables and the loose organisation of DAPS' overlay network it is
considered by is between structured and unstructured.

% TODO: MAYBE REVIEW THE EVALUATION
\begin{center}
\begin{tabular}{ccc}
\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
\hline
%
low &
%
low &
%
medium
\end{tabular}
\end{center}

@@

% \begin{figure}[ht]
% \centering
% \subfigure[Underlying network with edge costs.] {
%   \includegraphics[scale=0.4]{img/algorithms/narada}
%   \label{figure:narada:underlying}
% }\qquad\qquad
% \subfigure[Peer A sends a message to the rest using regular broadcast.] {
%   \includegraphics[scale=0.4]{img/algorithms/narada2}
%   \label{figure:narada:regu}
% }\qquad\qquad
% \subfigure[Peer A sends a message to the rest using end system multicast.] {
%   \includegraphics[scale=0.4]{img/algorithms/narada3}
%   \label{figure:narada:multicast}
% }
% \caption{A visualisation of the Narada protocol.}
% \label{figure:narada}
% \end{figure}

\emph{Narada} \cite{CRZ2000,CRSZ2001,CRSZ2002} is a generic
protocol for creating self-adapting overlay networks, that can achieve
application-layer multicast communication without requiring IP multicast
infrastructure at the network layer\footnote{
  IP multicast is the term reffering to the method of sending IP datagrams to a
  group of receivers in a single transmission. Even though the method is
  available some years now, IP multicast has not taken off as anticipated since
  many believe that it violates the stateless design of the current Internet.
  For this reason it is widely deployed in contained environments like
  enterprises (i.e. IPTV applications - distance learning, video conferencing),
  commercial stock exchanges, and multimedia content delivery networks.
}. Although it was not originally designed as a P2P system nor can be considered
scalable in the contemporary sense of the term, Narada has been pioneering
(maybe along with \emph{Scattercast} \cite{C2000}) in that
it was the first to consider the feasibility of overlay-based, application
layer, services over the Internet that can take into account bandwidth and
latency properties of the physical underlying infrastructure. Additionally,
authors realised the inefficiency caused by the topology mismatch problem and
tried to encounter it by building a richer connected graph, called a
\emph{mesh}, and building per source minimum spanning trees\footnote{
  ENarada \cite{LYL2008} used Gossip protocol for the
construction
}.
%as shown in Figure~\ref{figure:narada}.
They also keep the graph and the trees
dynamically updated while nodes continue to join and leave the network.
Moreover, the protocol tries to ease the physical link stress, the overall
resource usage and the relative delay among end systems. Unfortunately, the most
important limitation of Narada is that although it works reasonably well for
small groups, it does not scale well for larger networks, therefore it is not
suitable for, potentially very large, P2P file sharing application networks.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Landmark Binning (LM)}\label{sec:landmark_binning}
% \begin{figure*}
% \centering
%   \includegraphics[scale=0.4]{img/algorithms/landmark_binning}
% \caption{Example 2D coordinate overlay and a sample routing path from node D to
% (x,y).}
% \label{fig:landmark_binning}
% \end{figure*}

Landmark Binning \cite{RHMKS2002} is an approach to partition close
by nodes into bins based on their distance to well known anchor nodes across the
Internet.%, as shown in Figure~\ref{fig:landmark_binning}.
In order to detect
locality, nodes use network latency (i.e. RTT) as a measurement technique. The
network latency, even though not always accurate, is selected in this work due
to its non-intrusive, transparent and easy to apply behaviour. In order for the
binning to work, few well known anchor servers with known physical locations on
the Internet. Authors estimate that a number of 12 such servers can prove
sufficient for this. Every newly arriving node measures its distance from these
landmarks and unilaterally decides to join a specific bin based on these
results. Specifically, the node measures its round-trip time to each of the
landmarks and orders the values in a decreasing order. The ordering represents
a \emph{bin}, in the sence of close-by nodes having the same landmark ordering
and hance belong to the same bin. This means that a landmark system consisting
of $m$ such nodes results in $m!$ potentional different bins.The operation of
the method is independent of the model incorporated by the overlay network and
thus it can be applied with no significant changes to both structured and
unstructured P2P systems. \cite{RHMKS2002} has the detailed description of such
algorithm for both architectures. Landmark binning is also considered as a good
candidate for content distribution networks in particular. The major
disadvantage of landmark binning is to install and maintain landmark servers on
different autonomous systems, all over the world. A typical P2P network usually
has a couple of million nodes connected at any given time, thus rendering
landmark servers an important parameter when designing for large scales.  The
authors claim that latency estimation does not drain much from the network
resources but in case scalability problems arise, they claim the solution is
replacing single landmark servers with clusters of servers within the same
physical area. However, this approach does not reduce the possibility of
excessive network traffic flow through these landmark servers. One other
possible problem is the incorrect binning caused by the use of inaccurate
methods for delay measurement, since tracking down, for example, network
latencies is far from proven to be safe for location estimation.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Topology adaptation}
\begin{figure}[ht]
\centering
\subfigure[Ineffictient overlay topology.] {
  \includegraphics[scale=0.4]{img/pdf/topology-adaptation-before.pdf}
  \label{figure:topology-adaptation:before}
}\qquad\qquad
\subfigure[Efficient overlay topology after adaptation.] {
  \includegraphics[scale=0.4]{img/pdf/topology-adaptation-after.pdf}
  \label{figure:topology-adaptation:after}
}
\caption{A simple example of how an overlay topology can be adapted.}
\label{figure:topology-adaptation}
\end{figure}

The topology adaptation based protocols modify the topology of the P2P network
using various techniques. The two commonly used approaches that can
back such feature and were investigated in this section are creating
\emph{spanning tree}s using connection graphs, and creating \emph{clusters} of
physically close nodes.

Narada and subsequent algorithms like AOTO, LTM, SBO, try to solve the problem
by building a richer connected graph and forming minimum spanning trees over
this graph that can efficiently route messages among peers. Even though spanning
trees provide efficient query performance, their maintenance costs can be
proven a real blocker. That is why the original Narada algorithm did not have
any real topology adaptation mechanism meaning that it was designed for small
groups (therefore not scalable). Its followers try to overcome these obstacles
with clever tricks like forming minimum spanning trees for the two-hop
neighbours ($N^2$) of each node, partitioning the graph into two random groups
where each group is responsible of different tasks, or performing local
optimizations dynamically on the overlay graph. The advantage of building
minimum spanning trees is that they maintain the connectivity on the network in
an efficient manner, while still preserving the overall search scope. However,
their construction and the update costs, especially in dynamic, high-churn
environments, cause large additional traffic overhead to the underlying
network \cite{CRZ2000,CRSZ2001,CRSZ2002}.

The cluster based approaches, on the other hand, select to link physically
closer nodes with each other. T2MC, for example, uses traceroute logs and DDNO
uses domain names to cluster close by nodes. Unfortunatelly, commonly used
methods for proximity detection across Internet do not always return
reliable results, therefore mapping accuracy is not guaranteed. Moreover,
traceroute, is a heavy-weight utility to be frequently used on the
network, let alone the fact that many network equipment vendors do not allow
traceroute calls at all. The most problematic aspect of clusterting, though, is
its nature per se. Limited connectivity among the various local domains can
significantly shrink the search scope, negatively affecting the query response
time the P2P end-user experiences. Of course, techniques that try to round up
such edges and balance the efficiency of the clustering approach with improved
connectivity, have also been proposed, like DDNO, which attempts to address this
limitation by forcing half of each node's connections to be with other, randomly
selected, nodes over the network.

Concluding, topology adaptation ultimatelly targets on the reduction of the
average cost of the paths from one node to another. This way someone can deduce
that the overlay network is better fit on top of the physical, underlying one.
This deduction is fair in the general case but not always. For example an
adaptive algorithm can exchange a slow edge in the overlay with a faster one
thus reducing the average latency of the network communication. The pitfall here
is that this new virtual link can traverse a fast AS-to-AS link meaning that
even though is reduced message round-trip-time, it has more cost in terms of
communication network economics and management.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Forwarding optimisation}

\begin{figure}[ht]
\centering
\subfigure[Typical blind flooding.] {
\includegraphics[scale=0.25]{img/pdf/forwarding-optimisation-before.pdf}
  \label{figure:forwarding-optimisation:before}
}\qquad\qquad
\subfigure[A node can decide where to forward the messages.] {
  \includegraphics[scale=0.25]{img/pdf/forwarding-optimisation-after.pdf}
  \label{figure:forwarding-optimisation:after}
}
\caption{Forwarding optimisation in constraining message flooding.}
\label{figure:forwarding-optimisation}
\end{figure}

% Researchers initially started considering alternatives for services who
% implement one-to-many communication schemes, like IP multicast, on the
% application layer (i.e. Narada). Application level multicast protocols
% are generic protocols that can be applied to both P2P file sharing systems,
% as well as content distribution networks. Application layer multicast, however,
% is not as efficient as the one enabled by IP and still struggles to ease the
% difficulties imposed by the topology mismatch problem.

Forwarding optimisation can reduce the emitted messages flooding the overlay and
as a concequence the overall burden, network resources have to handle. In the
first realisation of the Gnutella protocol, each query request received by a
peer, was then forwarded to all of its neighbours. This was the source of
unecessary, sometimes redundant traffic. Forwarding-optimisation-based
approaches, propose intelligent forwarding link selection for a message's next
hop across the routing path. The selection criteria vary depending on the
algorithm, but they mainly use one or more statistical heuristics. These can
span from some obvious ones, like the latency of the link between the nodes or
the exposure of specific features, such as high processing bandwidth, high
capacity, or low latency, to more sophisticated alternatives that may even take
into account application level requirements such as the reliability or
friendship. Figure~\ref{figure:forwarding-optimisation} shows a simple example
of how forwarding optimisation can work.
Subfigure~\ref{figure:forwarding-optimisation:before} shows an algorithm that
floods the entire network for search of an object the existence of which is
denoted by black nodes. Alternativelly, a node can decide to forward its
messages not to every output link but to a specific or subset of its output
links. In
Subfigure~\ref{figure:forwarding-optimisation:after}, the node in the middle,
does not flood its neighbours. Instead it picks one of them. The dashed line on
the right, shows that, for this routing process, the algorithm has rendered two
potential forwarding paths as inefficient (e.g., the target nodes show
overloading signs). These appoaches have the advantage of enhancing the search
responsiveness and reducing the aggregate resource usage of the physical
network. On the other hand, they suffer from drastic reduction of the search
scope (in Subfigure~\ref{figure:forwarding-optimisation:after} the object on the
far right is not reached) thus limiting the scalability of the whole network.
Moreover they don't actually solve the topology mismatch problem since they
don't give any guarantees that overlay and underlying topologies are aligned
with each other let alone quantify the mismatch and try to alleviate. Like all
other approaches, forwarding based methods are commonly applied in conjunction
with other to improve the quality of the P2P systems.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Caching and replication}

\begin{figure}[ht]
\centering
\subfigure[Indexing can reduce the cost of the last hop.] {
\includegraphics[scale=0.3]{img/pdf/replication-index.pdf}
  \label{figure:replication:index}
}\qquad\qquad
\subfigure[Data replication allong the path of a successful query.] {
  \includegraphics[scale=0.3]{img/pdf/replication-data.pdf}
  \label{figure:replication:data}
}
\caption{Index and data replication strategies.}
\label{figure:replication}
\end{figure}

Caching is a widely used technique to exploit locality and minimise redundant
transfer of data, initially and successfully adopted by web and file server
application environments. Since peers in a P2P system also operate as
servers, it is intuitively expected that P2P file sharing systems can also
benefit from caching in improving performance and reducing overall resource
usage. However, the design of caches in this context is non-trivial compared to
the web-based caching. Exactly, because each node can act as both a server and a
client, two important issues have to be considered at design time. First, the
lifetime of a query is short, as the nodes join and leave frequently. Second,
the result of a single query string is not always the same, as it is dependent
on the source of the query, the TTL value set for the messages, the current
interconnection of peers and the high volatility of the environment. So, in
order to develop a successful caching system for a P2P architecture, these
parameters also have to be considered. P2P caching/replication can be applied
on two different levels, namely caching indices or pointers to data (see
Subfigure~\ref{figure:replication:index}) or caching
the data itself (see
Subfigure~\ref{figure:replication:data}). Successfull implementations have
already been developed in some commercial P2P systems, like KazaA, which uses
caching of indices on its super-peer level. Unfortunatelly, even though the
state of the art in P2P algorithms using caching methods reduce the resource
usage of the network, they also fail to consistently address typical problems
of a topologicaly mismatched environment such as duplication of messages.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Landmarking}\label{sec:landmark}

\begin{figure}[ht]
\centering
  \includegraphics[scale=0.4]{img/pdf/landmarking.pdf}
\caption{Landmark binning during node bootstrap.}
\label{figure:landmarking}
\end{figure}

In landmark-based algorithms, nodes use network delay (e.g. RTT) as a
distance measurement method to position themselves with respect to ``a priori''
known servers (the landmark servers) on the Internet. Landmark servers are used
by nodes to estimate their positions based on the intutive assumption that nodes
with similar distances to a set of landmarks, are physically close to each
other, as well, over the network. In Figure~\ref{figure:landmarking} the newlly
arriving peer, measures its distance to an array of landmark point denoted as
$Ln$ and and makes a sorted list of the peers, say in increasing order. Then in
order to choose the already participating peers with which to connect to,
compares its list to the list of the potential neighbour. Peers with similar
measured distances to the these landmark point are very likely to be close to
each other in the physical level as well.

Landmark based protocols has four important drawbacks. Firstly the network
delay is not a reliable distance estimation method. For example, based on the
load on the network the delay to certain nodes or networks can change from time
to time, which will eventually affect the distance measurements and wrong
measurements will lead to wrong estimated positions for the nodes, or incorrect
and non optimal clusterings of the nodes. Secondly, relying on predefined nodes
make the whole paradigm not fully distributed and the landmark system prone to
become a single point of failure. Moreover, the third drawback of using
landmark servers is the cost of installing and maintaining this kind of
infrastructure across the whole Internet and for all the different autonomous
system domains. As popular P2P file sharing applications usually have millions
of peers connected at any time, it is not false to assume that the network costs
of maintaining these landmark servers will be quite high. A possible solution to
the scalability problem of the static landmark servers is to use ordinary nodes
as dynamic landmarks once they estimate their own positions. Even though this
approach scales much better than static landmark servers, still the measurement
accuracy problem affects the overall performance of the system. Moreover,it can
be characterised as a rather coarse-grained approximation, therefore not
particularly well suited for detecting the correct positions of nodes within
close distance to each other.
